Inefficient Concessions and Mediation

### 2017-09-12 draft.tex
revise lemma and inefficiency / mediation theorems

\begin{lemma}
	Assuming that a separating equilibrium exists when concessions have material value and are efficient (where $e=1$), it is optimal to give efficient gifts in a separating equilibrium.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:}
The benefit of a gift appears on both sides of the IC for both types and thus cancels out in the low type IC $(peg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+peg_h )$ and in the high type IC:  $(pX_{TT}^h+(1-p)X_{FF}^h - g_h +peg_h \geq X_{FF}^h +peg_h)$.\footnote{Elaboration and explanation of the IC constraints is in the proof of Theorem~\ref{theorem:2}.} Therefore, receiving a gift where $e<1$  does not affect incentives to truth tell about the countries' types. However, the expected separation utilities under this game are now
\be
U_h= pX_{TT}+(1-p)X_{FF} - g_h+peg_h 
\ee
and
\be
U_l=X_{FF}+peg_h 
\ee
 $\frac{dU}{de}>0$ for both countries so it is never optimal have $e<1$.\footnote{This result does not depend on the benefit so it holds for any monotonic benefit function. It also holds for any monotone cost function.} \hfill $\blacksquare$

@@

\begin{theorem}
	Under some parameters, the optimal equilibrium uses inefficient concessions.
	\label{theorem:4}
\end{theorem}
 \emph{Proof}:   Because it is relatively complicated to calculate equilibrium concessions in the general case, a slight simplification of the general model will be made for proof of inefficient concessions. Restrict functional forms of $T(\cdot),W(\cdot)$ and $D(\cdot)$ so that each is a scalar multiplied by its arguments. Also assume that $s_1=s_2=m_1=m_2=1.$ Because of linearity, equilibrium optimization has high types choosing $\alpha=1$ (because a concession would only come from another high type in equilibrium) and low types choosing $\alpha=0$. Note that none of these simplifications is essential for the results to hold but we make them for ease of calculation and demonstration of the results. 
The low type's IC constraint binds so with $e=1$, $g^*$ can be calculated to be $g^*= \frac{p(T+D)}{1-pT}$.
Under an equilibrium with purely inefficient concessions $(e=0)$, the minimum separating concession would be $g'=p(T+D)$.

@@

\begin{theorem}
	A mediator can be used to eliminate inefficient concessions.
	\label{theorem:5}
\end{theorem}
\emph{Proof:}  The low type IC ($pg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+pg_h$) binds, so the equilibrium benchmark gift is $g^{*}=p(X_{FT}^l-X_{FF}^l)$ and high type utility is $U_h^*= pX_{TT}+(1-p)X_{FF} - (1-p)g^*$). Under $M$, the low type IC can bind without having high types give a concession to low types. The equilibrium separating gift $g^{*M}$ will be larger than $g^*$ in this case, specifically 
\be
g^{*M}= \frac{p(X_{FT}^l-X_{FF}^l)}{1-p}
\ee
High type utility under $M$ is 
\be
U_h^M= pX_{TT}+(1-p)X_{FF}
\ee
because high types exchange concessions of the same value. $U_h^M>U_h^*$. Note that no additional wealth is created by $M$, but less is transferred to low types, and welfare is defined as high-type utility.\footnote{Low types IR constraint still binds and $U_l=X_{FF}$} \hfill $\blacksquare$

>>>>

\begin{lemma}
	Assuming that a separating equilibrium exists when concessions have material value and are efficient (where $e=1$), it is optimal to give efficient concessions in a separating equilibrium.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:}
The benefit of a gift appears on both sides of the IC for both types and thus cancels out in the low type IC $(peg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+peg_h )$ and in the high type IC:  $(pX_{TT}^h+(1-p)X_{FF}^h - g_h +peg_h \geq X_{FF}^h +peg_h)$.\footnote{Elaboration and explanation of the IC constraints is in the proof of Theorem~\ref{theorem:2}.} Therefore, receiving a gift where $e<1$  does not affect incentives to truth tell about the countries' types. However, the expected separation utilities under this game are now
\be
U_h= pX_{TT}+(1-p)X_{FF} - g_h+peg_h 
\ee
and
\be
U_l=X_{FF}+peg_h 
\ee
 $\frac{dU}{de}>0$ for both countries so it is never optimal have $e<1$.\footnote{This result does not depend on the benefit so it holds for any monotonic benefit function. It also holds for any monotone cost function.} \hfill $\blacksquare$

@@

\begin{theorem}
		Assume the high types are not patient enough to separate without concessions and that concessions have material value that can be destroyed. There are parameter under which the optimal equilibrium is a separating equilibrium in which concessions are inefficient.
	\label{theorem:4}
\end{theorem}
 \emph{Proof}:   Because it is relatively complicated to calculate equilibrium concessions in the general case, a slight simplification of the general model will be made for proof of inefficient concessions. Restrict functional forms of $T(\cdot),W(\cdot)$ and $D(\cdot)$ so that each is a scalar multiplied by its arguments. Also assume that $s_1=s_2=m_1=m_2=1.$ Because of linearity, equilibrium optimization has high types choosing $\alpha=1$ (because a concession would only come from another high type in equilibrium) and low types choosing $\alpha=0$. Note that none of these simplifications is essential for the results to hold but we make them for ease of calculation and demonstration of the results. 
The low type's IC constraint binds so with $e=1$, $g^*$ can be calculated to be $g^*= \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D}$.
Under an equilibrium with purely inefficient concessions $(e=0)$, the minimum separating concession would be $g'=p(T+D)$.

@@

\begin{theorem}
	Assume that concessions have future material value.
	\begin{enumerate}
		\item[(a)] If future material value cannot be destroyed, a mediator can be used to expanded the parameter space over which peace can be achieved.
		\item[(b)] If future material value can be destroyed, a mediator can be used to eliminate inefficient concessions and improve welfare.
	\end{enumerate}
	\label{theorem:5}
\end{theorem}
\emph{Proof:}  The low type IC ($pg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+pg_h$) binds, so the equilibrium benchmark gift is $g^{*}=p(X_{FT}^l-X_{FF}^l)$ and high type utility is $U_h^*= pX_{TT}+(1-p)X_{FF} - (1-p)g^*$). Under $M$, the low type IC can bind without having high types give a concession to low types. The equilibrium separating gift $g^{*M}$ will be larger than $g^*$ in this case, specifically 
\be
g^{*M}= \frac{p(X_{FT}^l-X_{FF}^l)}{1-p}
\ee
High type utility under $M$ is 
\be
U_h^M= pX_{TT}+(1-p)X_{FF}
\ee
because high types exchange concessions of the same value. $U_h^M>U_h^*$. Note that no additional wealth is created by $M$, but less is transferred to low types, and welfare is defined as high-type utility.\footnote{Low types IR constraint still binds and $U_l=X_{FF}$} \hfill $\blacksquare$
